---
title: 'Project 123d: Walleye Management on the Lower Green and Colorado Rivers'
author:
  affiliation: Utah Division of Wildlife Resources
  name: Christopher Michaud
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  word_document: default
  html_document:
    fig_caption: no
    keep_md: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(flextable)
library(gt)
library(UCRBtools)
library(kableExtra)
library(sf)
library(mapview)
#library(BrailleR)

theme_set(theme_bw())
mapviewOptions(basemaps = c("Esri.WorldImagery", "OpenTopoMap", "CartoDB.Positron", "OpenStreetMap"))

`%!in%` <- Negate(`%in%`)

assets_path <- "C:/Users/cmichaud/OneDrive - DOI/Documents/Projects/udwr-mfo_data-mgt/Reports/Assets/"
#tusher <- "tusher_diversion.jpg"
co_efish <- "efboat-westwater1.jpg"
gr_we <- "walleye-carlos.jpg"
#co_sm <- "smallmouth-nick1.jpg"
```


```{r data-io, include=FALSE}

# Fetch final datasets 123d_2020.csv's 
files <- list.files(path = "./Reports/Rept-data/", 
                    pattern = "*.csv$",
                    full.names = TRUE)

dat <- map(files, read_csv)

dat_name <- list()
dat_name <- as.list(stringr::str_extract(files, "(?<=[0-9]{4}\\_[a-z]{2}\\-).*(?=\\.csv)"))

names(dat) <- dat_name


tmp_site <- dat$site %>% 
  mutate(across(starts_with("tm_"), ymd_hms),
         across(starts_with("tm_"), with_tz, "America/Denver"))

tmp_fish <- dat$fish %>% 
  mutate(across(starts_with("tm_"), ymd_hms),
         across(starts_with("tm_"), with_tz, "America/Denver"))

  

```

```{r spatial-io, include=FALSE}
# ----- Centerline (NHD) -----
ctr <- st_read("C:/Users/cmichaud/OneDrive - DOI/Documents/Data/Spatial/ucrb_spat_v2.sqlite",
               layer = "nhd_ctr_lns") %>%
  filter(nm_rvr %in% c("Green River", "Colorado River"))  

# ----- RMI points (NHD) -----
pt1 <- st_read("C:/Users/cmichaud/OneDrive - DOI/Documents/Data/Spatial/ucrb_spat_v2.sqlite",
               layer = "nhd_rmi_pts")

# ----- Subset points to individual reaches -----
gr_t <- pt1 %>%
  filter(cd_rvr == "GR" &
           between(rmi_nhd, 100, 129.5)) 

gr_g <- pt1 %>%
  filter(cd_rvr == "GR" &
           between(rmi_nhd, 129.5, 141.2))

co_w <- pt1 %>% 
  filter(cd_rvr == "CO" &
           between(rmi_nhd, 65, 118))

# ----- Crop centerline to reach extents -----
lww <- st_crop(ctr, co_w) %>% 
  mutate(reach = "Lower Westwater")

grey <- st_crop(ctr, gr_g) %>% 
  mutate(reach = "Lower Gray Canyon")

tusher <- st_crop(ctr, gr_t) %>% 
  mutate(reach = "Tusher Reach")

```

```{r}
# Define year 
yr = unique(year(tmp_site$tm_end))
```


```{r data-transformation}
# Modify site tbl
site <- tmp_site %>% 
  mutate(cd_rch = case_when(cd_rvr == "CO" ~ "LCO",
                              cd_rvr == "GR" &
                                between(rmi_start_bel, 100, 129.5) ~ "LGR",
                              cd_rvr == "GR" &
                                rmi_start_bel > 129.5 ~ "DESO"),
         nm_rch = case_when(cd_rch == "LCO" ~ "Lower Colorado River",
                              cd_rch == "DESO" ~ "Lower Gray Canyon",
                              cd_rch == "LGR" ~ "Tusher"),
         date = as.Date(tm_end),
         week = week(tm_end)) %>% 
  left_join(dat$water, by = c("cd_rvr", "date"))

site_rch <- site %>%
  split(.$cd_rch)

reach_tmp <- select(site, id_site, cd_rch, nm_rch, cd_rvr, nm_rvr)
```




```{r summary-dfs-site}

samp_stats <- list(
  max = ~ max(.x, na.rm = TRUE),
  min = ~ min(.x, na.rm = TRUE)
)

s_summary <- site %>%
  group_by(cd_rvr, nm_rvr, nm_rch, cd_rch) %>% 
  summarise(across(.cols = c(matches("rmi_"), date, discharge, temp), samp_stats),
            effort_hr = round(sum(amt_effort_sec, na.rm = TRUE)/3600, 1),
            sampl_days = length(unique(as.Date(tm_end))),
            .groups = "drop") %>% 
  mutate_if(is.Date, ~ format(.x, "%d %B %Y")) %>% 
#  mutate_if(is.numeric, ~ format(.x, big.mark = ",")) %>% 
  mutate(across(contains("rmi"), ~format(round(., 1), nsmall = 1)),
         across(contains("discharge"), ~format(., big.mark = ",")),
         temp_rng = paste(temp_min, temp_max, sep = " - "),
         dis_rng = paste(discharge_min, discharge_max, sep = " - "),
#         across(dis_rng, ~format(., big.mark = ",")),
         .keep = "unused") %>% 
  arrange(dmy(.$date_min))

s_summary_rch <- s_summary %>% 
  split(.$cd_rch)

```



```{r summary-dfs-water}

water_stats <- list(
  max = ~ max(.x, na.rm = TRUE),
  min = ~ min(.x, na.rm = TRUE),
  median = ~ median(.x, na.rm = TRUE),
  n = ~ n())


w_summary <- dat$water %>%
  group_by(cd_rvr) %>% 
  summarise(across(.cols = c(discharge), .fns = water_stats),
            dt_max_flow = format(date[which.max(discharge)], "%d %B %Y"),
            dt_min_flow = format(date[which.min(discharge)], "%d %B %Y"),
            .groups = "drop") %>% 
  mutate_if(is.numeric, ~ format(.x, big.mark = ",")) 

w_summary_rvr <- w_summary %>% 
  split(.$cd_rvr)

s_tmp <- site %>% 
  group_by(cd_rvr, cd_rch, nm_rch, date) %>% 
  summarise(amt_effort_sec = sum(amt_effort_sec, na.rm = TRUE),
            .groups = "drop") 

water <- dat$water %>% 
  left_join(s_tmp, by = c("date", "cd_rvr")) %>% 
  left_join(tbl_rvr, by = "cd_rvr") %>% 
  filter(date < as.Date(paste0(year(now()), "-10-01")))

```



```{r cpue-data}
# Create simple CPUE df
cpue <- site %>%
  left_join(dat$count, by = "id_site") %>% 
  mutate(wk = week(tm_end),
         mo = month(tm_end)) 
  # inner_join(select(tbl_spp, -nativity), by = c("cd_spp" = "cd_spp")) %>% # join common and sci names for display
  # inner_join(rch, by = c("cd_rch" = "cd_rch", "cd_rvr" = "cd_rvr")) %>% 
  # inner_join(select(site, site_id, rvr_temp)) %>% 
  # mutate(cd_rch = ifelse(cd_rch %in% c("WW", "LCO"), "LWW", cd_rch),
  #        rch_name = case_when(cd_rch == "LWW" ~ "Lower Westwater",
  #                             cd_rch == "DESO" ~ "Lower Gray Canyon",
  #                             cd_rch == "LGR" ~ "Tusher"))

# Aggregate CPUE to date 
cpue_dt <- cpue %>% 
  group_by(cd_rvr, cd_rch, cd_spp, date = as.Date(tm_end), discharge, temp) %>% 
  summarise(tot_effort = sum(amt_effort_sec, na.rm = TRUE) / 3600,
            n_fish = sum(n_fish, na.rm = TRUE),
            .groups = "drop") %>% 
  mutate(cpue_dt = n_fish/tot_effort)

```

```{r}
# Aggregate to reach and calculate total cpue and split reach Green River
cpue_rch <- cpue %>% 
  group_by(cd_rch, nm_rch, cn_spp, cd_spp) %>% 
  summarise(fish_ct = sum(n_fish),
            tot_effort = round(sum(amt_effort_sec)/3600,2),
           .groups = "drop") %>% 
  mutate(cpue = round(fish_ct/tot_effort, 2)) %>% 
  filter(fish_ct > 0) %>% 
  split(.$cd_rch)


```

```{r}
# Aggregate to reach and lifestage calculate CPUE
ls_cpue_tmp <- tmp_site %>% 
#  filter(cd_rvr == "CO") %>% 
  select(id_site, cd_rvr, nm_rvr, amt_effort_sec) %>% 
  inner_join(tmp_fish, by = "id_site") %>% 
  filter(cd_spp %in% c("SM", "NP", "WE") &
           !is.na(tot_length)) 

ls_cpue <- ls_cpue_tmp %>% 
  life_stage(specvar = cd_spp, lenvar = tot_length, piscivore = FALSE) %>% 
  group_by(id_site, cd_rvr, nm_rvr, cd_spp, cn_spp, amt_effort_sec, ls) %>% 
  summarise(n_fish = sum(n_fish), 
            .groups = "drop") %>% 
  complete(nesting(id_site, cd_rvr, nm_rvr, cd_spp, cn_spp, amt_effort_sec),
           nesting(ls), 
           fill = list(n_fish = 0)) %>% 
  mutate(cpue = (amt_effort_sec / 3600) * n_fish)
  # group_by(cd_rvr, nm_rvr, cd_spp, cn_spp,ls) %>% 
  # summarise(amt_effort_hr = sum(amt_effort_sec)/3600,
  #           n_fish = sum(n_fish),
  #           .groups = "drop") %>%
  # mutate(cpue = round(n_fish/amt_effort_hr, 2))
  
```



```{r biomet-data}

fish_bio <- left_join(dat$fish, reach_tmp, by = c("id_site")) %>%
  split(.$cd_rch)

```

```{r lf-figures}
fig_lf <- function(.data = fish_bio, .x, .fig_num, binwidth = 25) { 
  .data %>% 
    filter(cd_spp == .x) %>% 
    ggplot() +
    geom_histogram(aes(x = tot_length),
                   binwidth = binwidth,
                   color = "black") +
    labs(title = paste0("Figure ", .fig_num, ". Length frequency histogram for ", 
                       unique(.data$cn_spp[.data$cd_spp == .x]), 
                       "\n encountered within the ", 
                       unique(.data$nm_rch),
                       " Reach, ",
                       unique(year(.data$tm_encounter))),
         x = "Total length (mm)",
         y = "Number of fish")
}

# fish_bio$LWW %>%
#   fig_lf(.x = "WE", .fig_num = 0)
# 
# # fish_bio$LGR %>%
# #   fig_lf(.x = "LG", binwidth = 5)
# # 
# # fish_bio$LGR %>%
# #   fig_lf(.x = "SM")
# 
# rbind(fish_bio$LCO, fish_bio$WW) %>%
#   fig_lf(.x = "SM")
```

```{r water-figs}

  
hydro <- function(.data, .river_code, .fig_num) {
  .river_name <- unique(na.omit(.data$nm_rvr[.data$cd_rvr == .river_code]))
  .data2 <- .data %>% 
    filter(cd_rvr == .river_code)
  
  ggplot(.data2) +
    geom_line(aes(x = date, y = discharge), lwd = 1) +
    geom_point(data = filter(.data2, !is.na(cd_rch)),
               aes(x = date, y = discharge), 
               pch = 21,
               cex = 4,
               fill = "dodgerblue",
               alpha = .8) +
    labs(title = paste0("Figure ", .fig_num, ". Discharge on the ", 
                        unique(.data2$nm_rvr),
                        ", ",
                        max(year(.data2$date), na.rm = TRUE)),
         subtitle = paste("Sampling events noted with blue points"),
         x = "",
         y = "Discharge (cfs)")  
}
 

 # Test:
 # hydro(water, .river_code = "GR", .fig_num = 12)
```

# {.tabset .tabset-fade .tabset-pills}

### Abstract

Utah Division of Wildlife Resources (UDWR, Moab, UT) expended `r sum(as.numeric(s_summary$effort_hr))` hours of electrofishing effort under Project 123d to remove nonnative walleye and smallmouth bass from the upper Colorado River Basin in `r unique(year(site$tm_end))`. Crews focused removal efforts within specific subreaches where catch rates are highest: downstream of Tusher Diversion (River Mile (RM) `r s_summary_rch$LGR$rmi_start_bel_max` - `r s_summary_rch$LGR$rmi_end_bel_min`) on the Green River and below Westwater Canyon (RM `r s_summary_rch$LCO$rmi_start_bel_max` - `r s_summary_rch$LCO$rmi_end_bel_min`) on the Colorado River. A total of `r sum(cpue$n_fish[cpue$cd_spp == "WE"])` walleye and `r sum(cpue$n_fish[cpue$cd_spp == "SM"])` smallmouth bass were encountered throughout sampling in `r yr`. While walleye encounters were similar to recent years, smallmouth bass captures on the Lower Colorado River were substantially more frequent. <br><br>


```{r summary-flx}
s_summary %>%

  flextable(col_keys = c("nm_rch", "date_min", "date_max", 
                         "sampl_days", "effort_hr", "temp_rng", "dis_rng")) %>% 
  set_table_properties(width = .7, layout = "autofit") %>% 

  theme_vanilla() %>% 

  align(align = "center", part = "all") %>% 
  set_header_labels(
    nm_rch = "Reach",
    date_min = "Start Date",
    date_max = "End Date",
    sampl_days = "Days Sampled",
    effort_hr = "Sampling Effort (hr)",
    temp_rng = "Temperature Range (°C)",
    dis_rng = "Discharge Range (cfs)"
  ) %>% 

  bg(bg = "#D3D3D3", part = "header") %>% 
  add_header_lines(values = paste("Table 1. Effort summary by reach in", yr), top = TRUE) %>% 
  bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")

```

___

<br><br> 

<center>
```{r study-area}

# ----- Map it -----
# mapview(grey, color = "green3", 
#         lwd = 5,
#         layer.name = "Lower Gray Canyon") +
  mapview(tusher, color = "blue", 
          lwd = 5, 
          layer.name = "Tusher") +
  mapview(lww, color = "red3", 
          lwd = 5,
          layer.name = "Lower Colorado River")

```
</center>

Figure 1. Removal reaches sampled during 2021: Tusher Reach (RM `r s_summary_rch$LGR$rmi_start_bel_max` - `r s_summary_rch$LGR$rmi_end_bel_min`) and Lower Colorado Reach (RM `r s_summary_rch$LCO$rmi_start_bel_max` -`r s_summary_rch$LCO$rmi_end_bel_min`).

<br><br>


___


## Green River

<br><br>

<center>
```{r fig.cap="A photograph showing UDWR fisheries technician Carlos Montanez with 14 walleye dislpayed on the deck of an electrofishing boat. All fish are oriented heads left and tails right and have white zipties secruing identification tags throughh their gills. A scale, bucket and measuring board are also visible."}
knitr::include_graphics(path = paste0(assets_path, gr_we),
                        dpi = 75)
```
</center>  

UDWR fisheries technician Carlos Montanez with 14 walleye on the deck of an electrofishing boat. All fish were later frozen and transported to Fort Collins, CO for use in an energetics study. 

___

<br><br>

### Tusher Reach

The Green River peaked at `r w_summary_rvr$GR$discharge_max` cfs on `r w_summary_rvr$GR$dt_max_flow`, measured on USGS gage 09315000, located at the town of Green River, Utah (Figure 2). Crews began sampling below Tusher Diversion on `r s_summary_rch$LGR$date_min` and ended on `r s_summary_rch$LGR$date_max`. Main channel temperatures ranged between `r s_summary_rch$LGR$temp_rng` °C during sampling (Table 1). Crews sampled between river miles `r s_summary_rch$LGR$rmi_start_bel_max` and `r s_summary_rch$LGR$rmi_end_bel_min` (Figure 1).

<br><br>


```{r gr-dis-cap}
ck <- water %>% 
  filter(cd_rvr == "GR") %>% 
  distinct(date, .keep_all = TRUE)
h_gr <- hydro(ck, .river_code = "GR", .fig_num = 2)
cap1 <- paste(capture.output(BrailleR::VI(h_gr)), collapse = " ")
extra_text <- "This figure is depicts the dates on which sampling occurred (points) and the river discharge (line) during the sampling period. The sampling dates and river discharge follow -"
cap_dat <- filter(water, cd_rvr == "GR") %>% 
  distinct(date, .keep_all = TRUE)
dt_sample <- format(cap_dat$date[cap_dat$cd_rvr == "GR" & !is.na(cap_dat$cd_rch)], "%B %d")
dis_sample <- paste(cap_dat$discharge[cap_dat$cd_rvr == "GR" & !is.na(cap_dat$cd_rch)], "cfs")

dt_dis <- paste0(dt_sample, ": ", dis_sample)
fnl_dt_dis <- paste(dt_dis, collapse = ", ")

fnl_cap1 <- paste(c(cap1, extra_text, fnl_dt_dis), collapse = " ")

```

<center>
```{r gr-dis-fig, fig.cap=fnl_cap1}

#fig.cap=fnl_cap1, 
h_gr

```

</center>


A total of `r s_summary_rch$LGR$effort_hr` hours of effort were allocated to this reach and  `r cpue_rch$LGR$fish_ct[cpue_rch$LGR$cn_spp == "walleye"]` walleye were removed (catch-per-unit-effort [cpue] = `r cpue_rch$LGR$cpue[cpue_rch$LGR$cn_spp == "walleye"]` fish-per-hour).  Walleye catch rates are similar to recent years (Table 4), but with increased effort substantially more walleye have been removed in 2019, 2020 and 2021 than previous years. Walleye total lengths within this reach were also similar to recent years (mean = `r round(mean(fish_bio$LGR$tot_length[fish_bio$LGR$cd_spp == "WE"], na.rm = TRUE))` mm, range = `r min(fish_bio$LGR$tot_length[fish_bio$LGR$cd_spp == "WE"], na.rm = TRUE)` - `r max(fish_bio$LGR$tot_length[fish_bio$LGR$cd_spp == "WE"], na.rm = TRUE)` mm) (Figure 3). Additionally, researchers removed `r cpue_rch$LGR$fish_ct[cpue_rch$LGR$cn_spp == "smallmouth bass"]` smallmouth bass (cpue = `r cpue_rch$LGR$cpue[cpue_rch$LGR$cn_spp == "smallmouth bass"]` fish-per-hour) and `r cpue_rch$LGR$fish_ct[cpue_rch$LGR$cn_spp == "northern pike"]` northern pike (cpue = `r cpue_rch$LGR$cpue[cpue_rch$LGR$cn_spp == "northern pike"]` fish-per-hour) (Table 2). Native fish encounters are summarized in Table 3. 
  
<br><br>

<center>
```{r nnf-cpue-GR-gt}

cpue_rch$LGR %>% 
  filter(cd_spp %!in% c(spp_nat, "UI")) %>%
  arrange(desc(cpue)) %>% 
  select(-c(cd_rch)) %>% 
  
  flextable(col_keys = c("cn_spp", "fish_ct", "cpue")) %>% 
  set_table_properties(width = .5, layout = "autofit") %>% 
  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    cn_spp = "Common Name",
    fish_ct = "Number of Fish Encountered",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
  add_footer_lines(values = paste0("Based on ", 
                                  unique(cpue_rch$LGR$tot_effort), 
                                  " hours of electrofishing effort"), 
                   top = FALSE) %>%
  add_header_lines(values = paste("Table 2. Nonnative fish encounters within\n the Tusher Reach,", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")

```
</center>


<br><br>

```{r nat-cpue-GR}
cpue_rch$LGR %>% 
  filter(cd_spp %in% spp_nat) %>%
  arrange(desc(cpue)) %>% 
  select(-c(cd_rch)) %>% 

  flextable(col_keys = c("cn_spp", "fish_ct", "cpue")) %>% 
  set_table_properties(width = .5, layout = "autofit") %>% 

  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    cn_spp = "Common Name",
    fish_ct = "Number of Fish Encountered",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
  add_footer_lines(values = paste0("Based on ", 
                                  unique(cpue_rch$LGR$tot_effort), 
                                  " hours of electrofishing effort"), 
                   top = FALSE) %>% 
  add_header_lines(values = paste("Table 3. Native fish encounters within the\n Tusher Reach,", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")
```

<br><br>


```{r}
tibble(yr = c("2021", "2020", "2019", "2018", "2017"),
       n_we = c(101, 113, 151, 12, 32),
       tot_effort = c(31.3, 39.7, 43.4, 7.2, 25),
       cpue = c(3.22, 2.85, 3.48, 1.66, 1.28)) %>% 
  flextable() %>% 
    set_table_properties(width = .6, layout = "autofit") %>% 

  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    yr = "Year",
    n_we = "Number of Fish Encountered",
    tot_effort = "Total Effort (hours)",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
 
  add_header_lines(values = paste("Table 4. Walleye catch-per-unit-effort \nTusher Reach, 2017 -", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")
```

<br><br>

```{r lf-gr-cap}
lf_gr <- fig_lf(.data = fish_bio$LGR, .x = "WE", .fig_num = 3)

cap3 <- paste(capture.output(BrailleR::VI(lf_gr)), collapse = " ")

#generate alt data
sum_lf <- summary(fish_bio$LGR$tot_length[fish_bio$LGR$cd_spp == "WE"])
nm_sum_lf <- names(sum_lf)
vec_lf <- as.vector(round(sum_lf))
dat_lf <- paste(nm_sum_lf, " = ", vec_lf, collapse = ", ")

# better alt text
extra_text <- " The visualized data can be summarized as follows: "
fnl_cap3 <- paste(c(cap3, extra_text, dat_lf), collapse = "")
```

<center>
```{r lf-gr-fig, fig.cap=fnl_cap3}
lf_gr
```
</center>

___

<br><br>


## Colorado River

<br><br>

<center>
```{r fig.cap="A photograph showing a researcher on the front of an electrofishing boat, holding a dip net. They are wearing black pants and a blue shirt with their back to the camera. The river in front of the boat is an opaque tan color. The black shist and sandstone canyons of Westwater are visible in the background."}
knitr::include_graphics(path = paste0(assets_path, co_efish),
                        dpi = 200)
```
</center>

Electrofishing below Westwater Canyon.

___

<br><br>

### Lower Colorado Reach

The Colorado River, at USGS gage 09180500, located near Cisco, UT, peaked at `r w_summary_rvr$CO$discharge_max` cfs on `r w_summary_rvr$CO$dt_max_flow` (Figure 4).  Crews began sampling the lower Colorado reach on `r s_summary_rch$LCO$date_min` and ended on `r s_summary_rch$LCO$date_max`. Main channel temperatures ranged between `r s_summary_rch$LCO$temp_rng`°C during sampling (Table 1). Crews sampled between river miles `r s_summary_rch$LCO$rmi_start_bel_max` and`r s_summary_rch$LCO$rmi_end_bel_min` (Figure 1). UDWR began non-native fish removal work within this reach in 2014. Over the past three years researchers have documented an increasing trend in smallmouth bass encounters (Table 7). During the 2020 and 2021 field seasons UDWR applied additional effort, compared to past years, to combat this increasing smallmouth bass presence.


<br><br>


```{r}
h_co <- hydro(water, .river_code = "CO", .fig_num = 4)

cap3 <- paste(capture.output(BrailleR::VI(h_co)), collapse = " ")

# Modify caption to include data and richer text
extra_text <- "This figure is depicts the dates on which sampling occurred (points) and the river discharge (line) during the sampling period."
# cap_dat <- filter(water, cd_rvr == "CO" & !is.na(cd_rch)) %>% 
#   distinct(date, .keep_all = TRUE)
# dt_sample <- format(cap_dat$date[cap_dat$cd_rvr == "CO" & !is.na(cap_dat$cd_rch)], "%B %d")
# dis_sample <- paste(cap_dat$discharge[cap_dat$cd_rvr == "CO" & !is.na(cap_dat$cd_rch)], "cfs")
# 
# dt_dis <- paste0(dt_sample, ": ", dis_sample)
# fnl_dt_dis <- paste(dt_dis, collapse = ", ")

fnl_cap3 <- paste(c(cap3, extra_text), collapse = " ")
```

<center>
```{r fig.cap=fnl_cap3}
h_co
```

</center>

A total of `r s_summary_rch$LCO$effort_hr` hours of effort was allocated to this reach and crews removed `r cpue_rch$LCO$fish_ct[cpue_rch$LCO$cn_spp == "walleye"]` walleye (catch-per-unit-effort [cpue] = `r cpue_rch$LCO$cpue[cpue_rch$LCO$cn_spp == "walleye"]` fish-per-hour).  Additionally, `r cpue_rch$LCO$fish_ct[cpue_rch$LCO$cn_spp == "smallmouth bass"]` smallmouth bass were removed (cpue = `r cpue_rch$LCO$cpue[cpue_rch$LCO$cn_spp == "smallmouth bass"]`) within this reach (Table 5, Figure 5). This catch-rate is substantially higher than recent years (Table 7). This is particularly concerning due to the close proximity of the Westwater Canyon humpback chub population. Smallmouth bass mean total length in 2021 was `r round(mean(fish_bio$LCO$tot_length[fish_bio$LCO$cd_spp == "SM"], na.rm = TRUE))` mm (range = `r min(fish_bio$LCO$tot_length[fish_bio$LCO$cd_spp == "SM"], na.rm = TRUE)` - `r max(fish_bio$LCO$tot_length[fish_bio$LCO$cd_spp == "SM"], na.rm = TRUE)` mm) (Figure 6). Ancillary captures of native species are noted in Table 6. 


<br><br>


```{r nnf-cpue-CO}
cpue_rch$LCO %>% 
  filter(cd_spp %!in% spp_nat) %>%
  arrange(desc(cpue)) %>% 
  select(-c(cd_rch)) %>% 
  
  flextable(col_keys = c("cn_spp", "fish_ct", "cpue")) %>% 
  set_table_properties(width = .6, layout = "autofit") %>% 
  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    cn_spp = "Common Name",
    fish_ct = "Number of Fish Encountered",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
  add_footer_lines(values = paste0("Based on ", 
                                  unique(cpue_rch$LCO$tot_effort), 
                                  " hours of electrofishing effort"), 
                   top = FALSE) %>%
  add_header_lines(values = paste("Table 5. Nonnative fish encounters within the \nLower Colorado Reach,", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")
```

<br><br>

```{r nat-cpue-CO}
cpue_rch$LCO$cn_spp[cpue_rch$LCO$cd_spp == "CH"] <- "native Gila not identified to species"
cpue_rch$LCO %>% 
  filter(cd_spp %in% spp_nat) %>%
  arrange(desc(cpue)) %>% 
  select(-c(cd_rch)) %>% 

  flextable(col_keys = c("cn_spp", "fish_ct", "cpue")) %>% 
  set_table_properties(width = .6, layout = "autofit") %>% 

  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    cn_spp = "Common Name",
    fish_ct = "Number of Fish Encountered",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
  add_footer_lines(values = paste0("Based on ", 
                                  unique(cpue_rch$LCO$tot_effort), 
                                  " hours of electrofishing effort"), 
                   top = FALSE) %>% 
  add_header_lines(values = paste("Table 6. Native fish encounters within the \nLower Colorado Reach,", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")
```

<br><br>

```{r}
tibble(yr = c("2021", "2020", "2019", "2018", "2017"),
       n_sm = c(1164, 369, 12, 5, 35),
       tot_effort = c(43.5, 19.7, 6.6, 2, 7),
       cpue = c(26.78, 18.69, 1.81, 2.45, 4.96)) %>% 
  flextable() %>% 
    set_table_properties(width = .6, layout = "autofit") %>% 

  theme_vanilla() %>% 
  align(align = "center", part = "all") %>% 

  set_header_labels(
    yr = "Year",
    n_sm = "Number of Fish Encountered",
    tot_effort = "Total Effort (hours)",
    cpue = "Fish-per-Hour"
  ) %>% 
  bg(bg = "#D3D3D3", part = "header") %>%
 
  add_header_lines(values = paste("Table 7. Smallmouth bass catch-per-unit-effort \nLower Colorado Reach, 2017 -", yr), 
                   top = TRUE) %>% 
    bold(part = "header") %>% 
  fontsize(i = 1, size = 14, part = "header") %>% 
  align(j = 1, align = "left", part = "all")
```

<br><br>

```{r cpue-ls-cap, eval=TRUE, warning=FALSE}
cpue_ls_fig <- ls_cpue %>% 
  filter(cd_rvr == "CO" &
           cd_spp == "SM") %>% 
    mutate(ls_fct = as_factor(ls), 
         ls_fct = fct_recode(ls_fct,
                             Sub_Juvenile = "SUB-JUV",
                             Juvenile = "JUV",
                             Adult = "ADULT"),
         ls_fct = fct_relevel(ls_fct, "Sub_Juvenile", "Juvenile", "Adult")) %>%
  
  
  ggplot() +
  geom_boxplot(aes(x = ls_fct, y = cpue), fill = "grey") +
  scale_y_log10() +
  labs(title = "Figure 5. CPUE by size class for smallmouth bass \nencountered within the Lower Colorado Reach, 2021",
       x = "",
       y = "Fish per hour")

cap4 <- paste(capture.output(BrailleR::VI(cpue_ls_fig)), collapse = " ")
```

<center>
```{r cpue-ls-fig, fig.cap=cap4, eval=TRUE, warning=FALSE}
cpue_ls_fig

```
</center>

<br><br>


```{r eval=TRUE, warning=FALSE}
# Too few encounters for LFH!!!!

lf_co <- fish_bio$LCO %>% 
  fig_lf(.x = "SM", .fig_num = 6)
# fig_lf(.data = fish_bio$LCO, .x = "WE")
# fig_lf(.data = fish_bio$LCO, .x = "SM")
# fig_lf(.data = fish_bio$LGR, .x = "SM")

cap5 <- paste(capture.output(BrailleR::VI(lf_co)), collapse = " ")

#generate alt data
sum_lf <- summary(fish_bio$LCO$tot_length[fish_bio$LCO$cd_spp == "SM"])
nm_sum_lf <- names(sum_lf)
vec_lf <- as.vector(round(sum_lf))
dat_lf <- paste(nm_sum_lf, " = ", vec_lf, collapse = ", ")

# better alt text
extra_text <- " The visualized data can be summarized as follows: "
fnl_cap5 <- paste(c(cap5, extra_text, dat_lf), collapse = "")

```

<center>
```{r fig.cap=fnl_cap5, eval=TRUE, warning=FALSE}
lf_co
```

</center>

<br><br>

___


## Bureau of Reclamation Information

### COLORADO RIVER RECOVERY PROGRAM			RECOVERY PROGRAM
### FY `r yr` ANNUAL PROJECT REPORT				PROJECT NUMBER: <u> 123d </u>

I.	Project Title: Walleye control in the Lower Green and Colorado Rivers.

II.	Bureau of Reclamation Agreement Number(s): <br><br> 
  UDWR Moab: R19AP00059 <br>                                                 
  Project/Grant Period:  
  Start date (Mo/Day/Yr): 10/01/2019  
  End date: (Mo/Day/Yr): 09/30/2024  
  Reporting period end date: 09/30/2020  
  Is this the final report? Yes __ No  <u> X </u> <br><br>
  
  
III.	Principal Investigator(s): <br><br>
  Chris Michaud  
  Utah Division of Wildlife Resources  
  1165 S Hwy 191, Suite 4, 
  Moab, UT 84532  
  Phone: 435-259-3784;  
  E-mail: cmichaud@utah.gov <br><br>


IV.	Abstract: <br><br>
Utah Division of Wildlife Resources (UDWR, Moab, UT) expended `r sum(as.numeric(s_summary$effort_hr))` hours of electrofishing effort under Project 123d to remove nonnative walleye from the upper Colorado River Basin in `r unique(year(site$tm_end))`. Crews focused removal efforts within specific subreaches where walleye catch rates are highest: downstream of Tusher Diversion (River Mile (RM) `r s_summary_rch$LGR$rmi_start_bel_max` - `r s_summary_rch$LGR$rmi_end_bel_min`) on the Green River and below Westwater Canyon (RM `r s_summary_rch$LCO$rmi_start_bel_max` - `r s_summary_rch$LCO$rmi_end_bel_min`) on the Colorado River. A total of `r sum(cpue$n_fish[cpue$cd_spp == "WE"])` walleye and `r sum(cpue$n_fish[cpue$cd_spp == "SM"])` smallmouth bass were encountered throughout sampling in `r yr`. While walleye encounters were similar to recent years, smallmouth bass captures on the Lower Colorado River were substantially more frequent. <br><br>


V.	Study Schedule: <br><br>
2014-ongoing <br><br>

VI.	Relationship to RIPRAP: <br>  
GENERAL RECOVERY PROGRAM SUPPORT ACTION PLAN   
Reduce negative impacts of nonnative fishes and sportfish management 
activities (nonnative and sportfish management).  
Reduce negative interactions between nonnative and endangered fishes.  
Identify and implement viable active control measures. <br>  
GREEN RIVER ACTION PLAN: MAINSTEM  
Reduce impacts of nonnative fishes and sportfish management activities 
(nonnative and sportfish management).  
Reduce negative impacts to endangered fishes from sportfish management
activities.  
Develop and implement control programs for nonnative fishes in river reaches
occupied by the endangered fishes to identify required levels of control. Each
control activity will be evaluated for effectiveness, and then continued as needed. <br>  
COLORADO RIVER ACTION PLAN: MAINSTEM  
Reduce impacts of nonnative fishes and sportfish management activities
(nonnative and sportfish management).  
Reduce negative interactions between nonnative and endangered fishes.  
Develop basinwide aquatic management plan to reduce nonnative fish impacts
while providing sportfishing opportunities. <br><br>

VII.	Accomplishment of FY 2021 Tasks and Deliverables, Discussion of Initial Findings and Shortcomings: <br><br>
See reach specific analysis tabs <br><br>

VIII.	Additional noteworthy observations: <br><br>
See reach specific analysis tabs <br><br>

IX.	Recommendations: <br><br>  

    * **Tusher Reach (Green River)**  
  Continued application of effort between Tusher Diversion and Green River
  State Park (RM 128-120) as this segment has the highest catch rates for walleye in
  the reach. Removal should be conducted spring through early summer, or when justified by 
  noteworthy catch rates. <br><br> 

    * **Below Westwater Canyon (Colorado River)**  
  Continued application of effort between Big Hole camp and Fish Ford (RM 116-105.5). Removal should be 
  conducted spring through early summer, or when justified by noteworthy catch rates of walleye *or* 
  smallmouth bass. <br><br>

X.	Project Status: <br><br>
On track and ongoing <br><br>

XI.	FY 2021 Budget Status

	A.	Funds Provided: $53,697
	B.	Funds Expended: $53,697
	C.	Difference:  $0
	D.	Percent of the FY 2021 work completed, and projected costs to complete: 100%
	E.	Recovery Program funds spent for publication charges: $0  <br><br>

XII.	Status of Data Submission: <br><br> 
Data will be uploaded into STReaMs by the end of January 2022. <br><br>

XIII.	Signed:	<br>                                  		          
Principal Investigator: <u> Christopher Michaud </u>	&nbsp;&nbsp;&nbsp;&nbsp;	Date: <u> 11/21/2021 </u>

<br><br>
